A Mechanised Library for Generic (Co)Programming

### 2024-02-05 paper/paper.tex
missing word
add note about obj.magic

\hypersetup{
  linktocpage,
  colorlinks,
  citecolor=BlueViolet,
  filecolor=red,
  linkcolor=Blue,
  urlcolor=BrickRed
}
\newcommand{\mpav}[1]{\textcolor{red}{\textsc{Marco}: #1}}
\newcommand{\dcas}[1]{\textcolor{red}{\textsc{David}: #1}}
\newcommand{\proofcomment}[1]{\text{\{ #1 \}}}
\newenvironment{proofof}[1] {\begin{proof}[Proof of {#1}]}{\end{proof}}
\newcommand{\eqdef}{\stackrel{\mathrm{\Delta}}{=}}
\newcommand{\bnfeq}{\mathrel{::=}}
\newcommand{\defeq}{\triangleq}
\newcommand{\rul}[3]{\frac{#2}{#3}\;  {\textrulelabel{#1}}}
\newcommand{\den}[1]{\llbracket #1 \rrbracket}
\newcommand{\jud}[3]{#1 \vdash #2 : #3}
\newcommand{\bden}[1]{\llparenthesis#1 \rrparenthesis}
\newcommand{\bigslant}[2]{{\raisebox{.2em}{$#1$}\left/\raisebox{-.2em}{$#2$}\right.}}
\newcommand{\quotient}[2]{\bigslant{#1}{#2}}
\newcommand{\curry}{\Lambda}
\newcommand{\uncurry}{\begin{sideways}\begin{sideways}$\Lambda$\end{sideways}\end{sideways}}
\newcommand{\Bool}{\mathbb{B}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Nat}{\N}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Sets}{\mathbf{Set}}
\newcommand{\blacklater}{\blacktriangleright}
\newcommand{\tot}{\mathcal{S}}
\newcommand{\PSh}{\ensuremath{\textbf{PSh}(\omega)}}

@@

\begin{itemize}
  \item McBride devised a generalisation of Bove and Capretta's general
recursive function monad~\cite{McBride15, BoveC01}
  \item Guarded recursion for productive definitions and general recursion
~\cite{AtkeyM13, PaviottiMB15}
  \item PACO library~\cite{HurNDV13}
  \item Sized Types \cite{HughesPS96, abs-1202-3496}
  \item Copatterns \cite{AbelPTS13, AbelP16}
  \item IMPORTANT!!!!! Coq extraction from recursion
    schemes~\cite{larchey2022braga} \dcas{Check bibtex}
  \item Not that related: extraction for mutable arrays~\cite{SAKAGUCHI2020102372}
  \item Recursion schemes in Coq (but not hylos)~\cite{MurataE19}
  \item EVEN MORE IMPORTANT!!! 2023 POPL paper on divide and conquer code in
    Coq~\cite{AbreuDHJMS23}.  Some differences to point out: they require
    functional-extensionality axiom. We are axiom-free. Little discussion about
    extraction. They focus mostly on erasing termination proofs, but they do
    not discuss "what does their extracted code look like". Their
    representation of Functors leads me to think that they will produce loads
    of "Obj.magic", but we'd need to check this. 
\end{itemize}

@@

Another approach to certifying termination of nonstructurally recursive
functions is the use of \emph{sized types}. Sized types were introduced in
\textcite{HughesPS96} as a way to track/verify various properties of recursive
programs, including productivity termination. The core idea of sized types is
that types express bounds on the sizes of recursive data structures. This
approach has been used to express nonstructurally recursive algorithms in Agda:
\textcite{CopelloTB14} used sized types in a straightforward formulation of
mergesort. Such an approach has also been used in MiniAgda~\cite{abs-1202-3496}.

>>>>

\hypersetup{
  linktocpage,
  colorlinks,
  citecolor=BlueViolet,
  filecolor=red,
  linkcolor=Blue,
  urlcolor=BrickRed
}
\newcommand{\mpav}[1]{\textcolor{red}{\textsc{Marco}: #1}}
\newcommand{\dcas}[1]{\textcolor{red}{\textsc{David}: #1}}
\newcommand{\mvol}[1]{\textcolor{red}{\textsc{Michael}: #1}}
\newcommand{\proofcomment}[1]{\text{\{ #1 \}}}
\newenvironment{proofof}[1] {\begin{proof}[Proof of {#1}]}{\end{proof}}
\newcommand{\eqdef}{\stackrel{\mathrm{\Delta}}{=}}
\newcommand{\bnfeq}{\mathrel{::=}}
\newcommand{\defeq}{\triangleq}
\newcommand{\rul}[3]{\frac{#2}{#3}\;  {\textrulelabel{#1}}}
\newcommand{\den}[1]{\llbracket #1 \rrbracket}
\newcommand{\jud}[3]{#1 \vdash #2 : #3}
\newcommand{\bden}[1]{\llparenthesis#1 \rrparenthesis}
\newcommand{\bigslant}[2]{{\raisebox{.2em}{$#1$}\left/\raisebox{-.2em}{$#2$}\right.}}
\newcommand{\quotient}[2]{\bigslant{#1}{#2}}
\newcommand{\curry}{\Lambda}
\newcommand{\uncurry}{\begin{sideways}\begin{sideways}$\Lambda$\end{sideways}\end{sideways}}
\newcommand{\Bool}{\mathbb{B}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Nat}{\N}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Sets}{\mathbf{Set}}
\newcommand{\blacklater}{\blacktriangleright}
\newcommand{\tot}{\mathcal{S}}
\newcommand{\PSh}{\ensuremath{\textbf{PSh}(\omega)}}

@@

\begin{itemize}
  \item McBride devised a generalisation of Bove and Capretta's general
recursive function monad~\cite{McBride15, BoveC01}
  \item Guarded recursion for productive definitions and general recursion
~\cite{AtkeyM13, PaviottiMB15}
  \item PACO library~\cite{HurNDV13}
  \item Sized Types \cite{HughesPS96, abs-1202-3496}
  \item Copatterns \cite{AbelPTS13, AbelP16}
  \item IMPORTANT!!!!! Coq extraction from recursion
    schemes~\cite{larchey2022braga} \dcas{Check bibtex}
  \item Not that related: extraction for mutable arrays~\cite{SAKAGUCHI2020102372}
  \item Recursion schemes in Coq (but not hylos)~\cite{MurataE19}
  \item EVEN MORE IMPORTANT!!! 2023 POPL paper on divide and conquer code in
    Coq~\cite{AbreuDHJMS23}.  Some differences to point out: they require
    functional-extensionality axiom. We are axiom-free. Little discussion about
    extraction. They focus mostly on erasing termination proofs, but they do
    not discuss "what does their extracted code look like". Their
    representation of Functors leads me to think that they will produce loads
    of "Obj.magic", but we'd need to check this. \mvol{We have a story for this now: obj.magic harms interaction with other ocaml code, the resulting ocaml code is not natural to read, and in general more complex/strange extracted code is likely to be less efficient and harder to optimise.}
\end{itemize}

@@

Another approach to certifying termination of nonstructurally recursive
functions is the use of \emph{sized types}. Sized types were introduced in
\textcite{HughesPS96} as a way to track/verify various properties of recursive
programs, including productivity and termination. The core idea of sized types
is that types express bounds on the sizes of recursive data structures. This
approach has been used to express nonstructurally recursive algorithms in Agda:
\textcite{CopelloTB14} used sized types in a straightforward formulation of
mergesort. Such an approach has also been used in MiniAgda~\cite{abs-1202-3496}.
